---
title: "Emergency Broadband Benefit analysis"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(easypackages)
libraries("readxl","ggplot2","stringr", "rlist", "data.table", "DBI", "RSQLite",  
          "stargazer", "Hmisc","magrittr","formattable", "RPostgreSQL", "dplyr", 
          "tidyverse", "dbplyr", "zipcodeR")
options(scipen=999)

```

Let's dig into the Emergency Broadband Benefit data, [located here](https://www.usac.org/about/emergency-broadband-benefit-program/emergency-broadband-benefit-program-enrollments-and-claims-tracker/additional-ebb-program-data/). 


```{r }

ebb_data <- read_xlsx("EBB_Enrollment_10_1_zip5.xlsx")

```

We need to convert the data from zip code to Census tract because it is so much easier. 

```{r }

download_zip_data(force = FALSE)

```







https://www.fcc.gov/reports-research/maps/tract-level-residential-fixed-connections-dec-2018/

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.




The FCC also provides the [Residential Fixed Internet Access Service Connections per 1000 Households by Census Tract](https://www.fcc.gov/reports-research/maps/tract-level-residential-fixed-connections-dec-2018/). 

```{r }

service_connections <- read.csv("tract_map_dec_2018.csv")

```
